Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-03-12 pkg/disttree/inst/draft/draft.Rnw
small corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2052 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 231a95db8ddaca3a64bf17d59717eb05b4c68937

With the specification of the distribution family and its log-likelihood 
function the task of fitting a distributional model turns into the task 
of determining the distribution parameter~$\theta$. This is done by the 
following maximization.
$$
\hat{\theta} = \max_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)
$$
where $\{y_i\}_{i=1,\ldots,n}$ are the observations of the response 
variable $Y$.

In that way, a whole distribution is specified with all its features, 
including location, scale, and shape. However, fitting one global model 
to the whole data set might be too much of a generalization and therefore 
not represent the data and its features in a reasonable way. Moreover, 
if covariates are available it is desirable to include them into the model 
as they might provide important additional information. In our application 
on precipitation data a wide range of covariates is available from a numerical 
ensemble prediction system considering precipitation, temperature, sea level 
pressure and other meteorological quantities which might be of interest and 
have an influence on the predictions.

@@

\subsection{Distributional tree}
\label{disttree}
Given the fit of a global distributional model $D(Y, \bold{\hat{\theta}})$ 
the idea is to use the information provided by this model together with a 
set of covariates $\bold{Z} = (Z_1, \ldots, Z_m) \in \mathcal{Z}$ 
to decide whether the data set should be split into subgroups or not. 
In particular, the information provided by the fitted model are the scores 
$$
s(\hat{\theta}, y_i) = \frac{\partial \ell}{\partial \theta}(\hat{\theta}; y_i).
$$ 
These values express how well the model fits the data regarding each 
parameter. Ideally they should fluctuate randomly around zero, 
similar to residuals in ordinary least squares estimation, but in 
this case there is a score value for each pair of estimated parameter 
$\hat{\theta}_j \in (\hat{\theta}_1, \ldots, \hat{\theta}_k) = \bold{\hat{\theta}}$ 
and observation $y_j \in \{y_i\}_{i = 1,\ldots, n}$. This enables the 
model to be sensitive to changes in each of the parameters which is one 
of the main advantages of distributional trees over other tree models 
considering only the mean of the distribution.

>>>>

With the specification of the distribution family and its log-likelihood 
function the task of fitting a distributional model turns into the task 
of determining the distribution parameter~$\theta$. This is done by the 
following maximization based on a learning sample $\{y_i\}_{i=1,\ldots,n}$ 
of the response variable $Y$.
$$
\hat{\theta} = \max_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)
$$
which is equal to solving 
$$
\sum_{i=1}^n \frac{\partial \ell}{\partial \theta}(\hat{\theta}; y_i) = 0.
$$
In that way, a whole distribution is specified with all its features, 
including location, scale, and shape. However, fitting one global model 
to the whole data set might be too much of a generalization and therefore 
not represent the data and its features in a reasonable way. Moreover, 
if covariates are available it is desirable to include them into the model 
as they might provide important additional information. In our application 
on precipitation data a wide range of covariates is available from a numerical 
ensemble prediction system considering precipitation, temperature, sea level 
pressure and other meteorological quantities which might be of interest and 
have an influence on the predictions.

@@

\subsection{Distributional tree}
\label{disttree}
Given the fit of a global distributional model $D(Y, \bold{\hat{\theta}})$, 
the idea is to use the information provided by this model together with a 
set of covariates $\bold{Z} = (Z_1, \ldots, Z_m) \in \mathcal{Z}$ 
to decide whether the data set should be split into subgroups or not. 
In particular, the information provided by the fitted model are the scores 
$$
s(\hat{\theta}, y_i) = \frac{\partial \ell}{\partial \theta}(\hat{\theta}; y_i).
$$ 
These values express how well the model fits the data regarding each 
parameter. Ideally they should fluctuate randomly around zero, 
similar to residuals in ordinary least squares estimation, but in 
this case there is a score value for each pair of estimated parameter 
$\hat{\theta}_j \in (\hat{\theta}_1, \ldots, \hat{\theta}_k) = \bold{\hat{\theta}}$ 
and observation $y_j \in \{y_i\}_{i = 1,\ldots, n}$. This enables the 
model to be sensitive to changes in each of the parameters which is one 
of the main advantages of distributional trees over other tree models 
considering only the mean of the distribution.
